---
title: "Statistical Methods for Discrete Response, Time Series, and Panel Data (W271): Lab 2"
author: "Harry Xu and Rich Ung"
output: 
  pdf_document:
  toc: true
  number_sections: true
fontsize: 11pt
geometry: margin=1in
line_spacing: single
---

# Introduction

## Strategic Placement of Products in Grocery Stores

We receive the following prompt from **Question 12 of chapter 3 (on page 189 and 190)** within Bilder and Loughin's *"Analysis of Categorical Data with R"*:

```In order to maximize sales, items within grocery stores are strategically placed to draw customer attention. This exercise examines one type of item—breakfast cereal. Typically, in large grocery stores, boxes of cereal are placed on sets of shelves located on one side of the aisle. By placing particular boxes of cereals on specific shelves, grocery stores may better attract customers to them. To investigate this further, a random sample of size 10 was taken from each of four shelves at a Dillons grocery store in Manhattan, KS. These data are given in the cereal_dillons.csv file. The response variable is the shelf number, which is numbered from bottom (1) to top (4), and the explanatory variables are the sugar, fat, and sodium content of the cereals.```

By using multicategory response models, we can maximize sales by placing the cereal in the shelf that best fits its attributes (such as sugar, fat, and sodium content).

# Exploratory Data Analysis

In order to perform our analyses, we first load the required R libraries:

```{r setup, message=FALSE, warning=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=60), tidy=TRUE)
library(package = MASS)
library(package = car)
library(package = nnet)
```

Next, we load the data and perform a quick EDA:

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
cereal <- read.csv('cereal_dillons.csv', header=TRUE)
head(cereal)
str(cereal)
summary(cereal)
```

We can see that our dataset has 40 observations of 7 variables, with no missing values for any of our variables. It looks like each row corresponds to a particular cereal, where "Cereal" contains the name of the cereal and "Shelf" contains the shelf that the cereal is located. We can also see that we have the serving size, sugar, fat, and sodium contents for each cereal.

# Modeling & Questions

We go through our modeling as we go through parts a through h within this question:

## Part A

#### Question

The explanatory variables need to be re-formatted before proceeding further. First, divide each explanatory variable by its serving size to account for the different serving sizes among the cereals. Second, re-scale each variable to be within 0 and $1.^{12}$ Below is code we use to re-format the data after the data file is read into an object named `cereal`:

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
stand01 <- function(x) { (x - min(x))/(max(x) - min(x)) }
cereal2 <- data.frame(Shelf = cereal$Shelf, sugar = 
                        stand01(x = cereal$sugar_g/cereal$size_g), fat =
                        stand01(x = cereal$fat_g/cereal$size_g), sodium =
                        stand01(x = cereal$sodium_mg/cereal$size_g))
```

#### Answer

By executing the code chunk above, we have already re-formatted the dataset by dividing each explanatory variable by its serving size to account for the different serving sizes among the cereals, and re-scaling each variable to be between 0 and 1 in order to help with the convergence of parameter estimates.

## Part B

#### Question

Construct side-by-side box plots with dot plots overlaid for each of the explanatory variables. Below is code that can be used for plots involving sugar:

```{r, eval = FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
boxplot(formula = sugar ~ Shelf, data = cereal2, ylab = "Sugar", xlab = "Shelf", pars = list(outpch=NA))
stripchart(x = cereal2$sugar ~ cereal2$Shelf, lwd = 2, col = "red", method = "jitter", vertical = TRUE, pch = 1, add = TRUE)
```

Also, construct a parallel coordinates plot for the explanatory variables and the shelf number. Discuss if possible content differences exist among the shelves.

#### Answer

Below are the side-by-side box plots with dot plots overlaid for each of the explanatory variables:

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60), fig.height = 4.5, fig.width = 5.5, fig.align = "center"}
boxplot(formula = sugar ~ Shelf, data = cereal2, ylab = "Sugar", xlab = "Shelf", pars = list(outpch=NA))
stripchart(x = cereal2$sugar ~ cereal2$Shelf, lwd = 2, col = "red", method = "jitter", vertical = TRUE, pch = 1, add = TRUE)

boxplot(formula = fat ~ Shelf, data = cereal2, ylab = "Fat", xlab = "Shelf", pars = list(outpch=NA))
stripchart(x = cereal2$fat ~ cereal2$Shelf, lwd = 2, col = "red", method = "jitter", vertical = TRUE, pch = 1, add = TRUE)

boxplot(formula = sodium ~ Shelf, data = cereal2, ylab = "Sodium", xlab = "Shelf", pars = list(outpch=NA))
stripchart(x = cereal2$sodium ~ cereal2$Shelf, lwd = 2, col = "red", method = "jitter", vertical = TRUE, pch = 1, add = TRUE)
```

Below is a parallel coordinates plot for the explanatory variables and the shelf number:

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60), fig.height = 4.5, fig.width = 5.5, fig.align = "center"}

cereal2.colors<-ifelse(test = cereal2$Shelf==1, yes = "black", 
                    no = ifelse(test = cereal2$Shelf==2, yes = "red",
                                ifelse(test = cereal2$Shelf==3, yes = "red", no = "blue")))

cereal2.lty<-ifelse(test = cereal2$Shelf==1, yes = "solid", 
                    no = ifelse(test = cereal2$Shelf==2, yes = "longdash",
                                ifelse(test = cereal2$Shelf==3, yes = "dotdash", no = "twodash")))

parcoord(x = cereal2, col = cereal2.colors, lty = cereal2.lty)  # Plot
legend(x = 3.5, y = 1, legend = c("1", "2", "3", "4"), lty = c("solid", "longdash", "dotdash", "twodash"),
      col=c("black", "red", "green", "blue"), cex=0.8, bty="n")

```

There appears to be some content difference clustered by shelves. As the above parallel coordinates plot shows, Shelf 4 Cereals appear to have the lowest Sugar, Fat and Sodium content -- which suggest this shelf contains "healthier" types of cereals. Shelf 1 appears to have the highest Sugar content and perhaps the most "unhealthiest" cereals. This appears to be the case by examining the brands of the Cereals placed on Shelf 4 vs. brands of Cereals placed on Shelf 1

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
cereal[cereal$Shelf==4,]
cereal[cereal$Shelf==1,]
```

## Part C

#### Question

The response has values of 1, 2, 3, and 4. Under what setting would it be desirable to take into account ordinality. Do you think this occurs here?

#### Answer

It would be desirable to take into account ordinality when the variable has a natural ordering to their levels. In other words, if response levels can be arranged so that category 1 < category 2 < ··· < category J in some conceptual scale of measurement (e.g., amount of agreement). Since the shelf has a natural ordering to their levels, bottom (1) to top (4), it would make sense to take into account ordinality.

## Part D

#### Question

Estimate a multinomial regression model with linear forms of the sugar, fat, and sodium variables. Perform LRTs to examine the importance of each explanatory variable.

#### Answer

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
mod.fit.ord <- polr(formula = as.factor(Shelf) ~ sugar + fat + sodium, data = cereal2, method = "logistic")
summary(mod.fit.ord)
Anova(mod.fit.ord)
```

The results of the LRT for the multionmial regression model show that the only significant variable appears to be Sodium in predicting which shelf the Cereal will be placed on. 

## Part E

#### Question

Show that there are no significant interactions among the explanatory variables (including an interaction among all three variables).

#### Answer

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
mod.fit.ord2 <- polr(formula = as.factor(Shelf) ~ sugar + fat + sodium
                     + sugar:fat + sugar:sodium + fat:sodium
                     + sugar:fat:sodium
                     , data = cereal2, method = "logistic")
Anova(mod.fit.ord2)
```

## Part F

#### Question

Kellogg’s Apple Jacks (http://www.applejacks.com) is a cereal marketed to- ward children. For a serving size of 28 grams, its sugar content is 12 grams, fat content is 0.5 grams, and sodium content is 130 milligrams. Estimate the shelf probabilities for Apple Jacks.

#### Answer

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
summary(mod.fit.ord)
predict(object = mod.fit.ord, newdata = data.frame(sugar = (((12/28) - min(cereal$sugar_g))/(max(cereal$sugar_g) - min(cereal$sugar_g))),
                                                   fat = (((0.5/28) - min(cereal$fat_g))/(max(cereal$fat_g) - min(cereal$fat_g))),
                                                   sodium = (((130/28) - min(cereal$sodium_mg))/(max(cereal$sodium_mg) - min(cereal$sodium_mg)))),  type = "probs")
```

## Part G

#### Question

Construct a plot similar to Figure 3.3 where the estimated probability for a shelf is on the y-axis and the sugar content is on the x-axis. Use the mean overall fat and sodium content as the corresponding variable values in the model. Interpret the plot with respect to sugar content.

#### Answer

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60), fig.height = 4.5, fig.width = 5.5, fig.align = "center"}
mod.fit.sugar<-polr(formula = as.factor(Shelf) ~ sugar + sodium + fat, data = cereal2, method = "logistic")
summary(mod.fit.sugar)
beta.hat<-c(-mod.fit.ord$coefficients, mod.fit.ord$zeta)

curve( 1/(1 + exp(beta.hat[3] + beta.hat[1]*x) + exp(beta.hat[4] + beta.hat[1]*x)),
      ylab = expression(hat(pi)), xlab = "Sugar",
      ylim = c(0,0.5),
      xlim = c(min(cereal2$sugar), max(cereal2$sugar)), 
      col = "black", lty = "solid", lwd = 2, n = 1000, type = "n",
      panel.first = grid(col = "gray", lty = "dotted"))

# Shelf 1
curve(expr = plogis(q = mod.fit.sugar$zeta[1] - mod.fit.sugar$coefficients[1]*x 
                    - mod.fit.sugar$coefficients[2]*mean(cereal2$sodium)
                    - mod.fit.sugar$coefficients[3]*mean(cereal2$fat)), 
      col = "green", type = "l", add = TRUE, lty = "dotdash", n = 1000)  

# Shelf 2
curve(expr = plogis(q = mod.fit.sugar$zeta[2] - mod.fit.sugar$coefficients[1]*x
                        - mod.fit.sugar$coefficients[2]*mean(cereal2$sodium)
                        - mod.fit.sugar$coefficients[3]*mean(cereal2$fat)) 
          - plogis(q = mod.fit.sugar$zeta[1] - mod.fit.sugar$coefficients[1]*x 
                    - mod.fit.sugar$coefficients[2]*mean(cereal2$sodium)
                    - mod.fit.sugar$coefficients[3]*mean(cereal2$fat)), 
      col = "red", type = "l", add = TRUE, lty = "longdash", n = 1000)  

# Shelf 3
curve(expr = plogis(q = mod.fit.sugar$zeta[3] - mod.fit.sugar$coefficients[1]*x
                        - mod.fit.sugar$coefficients[2]*mean(cereal2$sodium)
                        - mod.fit.sugar$coefficients[3]*mean(cereal2$fat)) 
          - plogis(q = mod.fit.sugar$zeta[2] - mod.fit.sugar$coefficients[1]*x 
                    - mod.fit.sugar$coefficients[2]*mean(cereal2$sodium)
                    - mod.fit.sugar$coefficients[3]*mean(cereal2$fat)), 
      col = "blue", type = "l", add = TRUE, lty = "solid", n = 1000)  

# Shelf 4
curve(expr = 1 - plogis(q=mod.fit.sugar$zeta[3]- mod.fit.sugar$coefficients[1]*x
                        - mod.fit.sugar$coefficients[2]*mean(cereal2$sodium)
                        - mod.fit.sugar$coefficients[3]*mean(cereal2$fat)),
     col='black', type="l", add = TRUE, lty = "twodash", n = 1000)

legend(x = 'bottomleft', legend=c("Shelf 1", "Shelf 2", "Shelf 3", "Shelf 4"), 
       lty=c("dotdash","longdash","solid", "twodash"),
       col=c("green","red","blue", "black"), bty="n", lwd = c(2,2,2,2), seg.len = 4)
```

By holding the Sodium and Fat variables constant at their mean values, we can plot the probability of being on each Shelf (1-4) relative to the explanatory variable Sugar. As we noted before, it appears that Shelf 4 contains the "healthy" cereals whereas Self 1 contains the "unhealthy" cereals. We can see that in the above probability curves since as the Sugar variable increases, the probability of being on Shelf 4 decreases while the probability of being on Shelf 1 increases.

## Part H

#### Question

Estimate odds ratios and calculate corresponding confidence intervals for each explanatory variable. Relate your interpretations back to the plots constructed for this exercise.

#### Answer

We convert c to be 1g or 1mg of the content in order to better explain the interpretation:

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

sd.cereal<-round(apply(X = cereal[,-c(1,2,3,4)], MARGIN = 2, FUN = sd))

sd.cereal

sugar_1g_conv <- (((sd.cereal[1]/28) - min(cereal$sugar_g))/(max(cereal$sugar_g) - min(cereal$sugar_g)))

fat_1g_conv <- (((sd.cereal[2]/28) - min(cereal$fat_g))/(max(cereal$fat_g) - min(cereal$fat_g)))

sodium_1mg_conv <- (((sd.cereal[3]/28) - min(cereal$sodium_mg))/(max(cereal$sodium_mg) - min(cereal$sodium_mg)))

c.value <- c(sugar_1g_conv, fat_1g_conv, sodium_1mg_conv)

round(exp(c.value * (-mod.fit.ord$coefficients)),2)

conf.beta <- confint(object = mod.fit.ord, level = 0.95)

ci <- exp(c.value*(-conf.beta))

round(data.frame(low = ci[,2], up = ci[,1]), 2)

summary(mod.fit.ord)

Anova(mod.fit.ord)

```

# Conclusion

In conclusion, we were able to predict...






